---
title: 'Homework 4: Machine Learning'
author: "Your name goes here"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r}
#| label: load-libraries
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false
options(scipen = 999) #disable scientific notation
library(tidyverse)
library(tidymodels)
library(GGally)
library(sf)
library(leaflet)
library(janitor)
library(rpart.plot)
library(here)
library(scales)
library(vip)
```


# The Bechdel Test


```{r read_data}

bechdel <- read_csv(here::here("data", "bechdel.csv")) %>% 
  mutate(test = factor(test)) 
glimpse(bechdel)

```
How many films fail/pass the test, both as a number and as a %?

```{r}
bechdel %>% 
  group_by(test) %>% 
  summarise('n'=n()) %>% 
  mutate("percent"=n/sum(n))
```
622 movies pass the test 44.62%
772 movies fail the test 55.38%

## Movie scores
```{r}
ggplot(data = bechdel, aes(
  x = metascore,
  y = imdb_rating,
  colour = test
)) +
  geom_point(alpha = .3, size = 3) +
  scale_colour_manual(values = c("tomato", "olivedrab")) +
  labs(
    x = "Metacritic score",
    y = "IMDB rating",
    colour = "Bechdel test"
  ) +
 theme_light()
```


# Split the data
```{r}
# **Split the data**

set.seed(123)

data_split <- initial_split(bechdel, # updated data
                           prop = 0.8, 
                           strata = test)

bechdel_train <- training(data_split) 
bechdel_test <- testing(data_split)
```

Check the counts and % (proportions) of the `test` variable in each set.
```{r}
bechdel_train %>% 
  group_by(test) %>% 
  summarise('n'=n()) %>% 
  mutate("percent"=n/sum(n))
bechdel_test %>% 
  group_by(test) %>% 
  summarise('n'=n()) %>% 
  mutate("percent"=n/sum(n))

```
The counts and proportions are as expected. The count is higher on the training data set but the proportions are the same (use of stata=test)

## Feature exploration

## Any outliers? 

```{r}

bechdel %>% 
  select(test, budget_2013, domgross_2013, intgross_2013, imdb_rating, metascore) %>% 

    pivot_longer(cols = 2:6,
               names_to = "feature",
               values_to = "value") %>% 
  ggplot()+
  aes(x=test, y = value, fill = test)+
  coord_flip()+
  geom_boxplot()+
  facet_wrap(~feature, scales = "free")+
  theme_bw()+
  theme(legend.position = "none")+
  labs(x=NULL,y = NULL)

```
From the above plot we can identify that there are some outliers. More specifically when it comes to budget there is one movie that failed the test with a huge budget. Also intgross has some extreme values for both categories. Regarding metascore there is an outlier with a vary small value for a movie that passed the test.
Based on the plot we can see that the budget for movies that pass the test is usually smaller, as well as the imdb rating.
## Scatterplot - Correlation Matrix

Write a paragraph discussing the output of the following 
```{r, warning=FALSE, message=FALSE}
bechdel %>% 
  select(test, budget_2013, domgross_2013, intgross_2013, imdb_rating, metascore)%>% 
  ggpairs(aes(colour=test), alpha=0.2)+
  theme_bw()
```

the distributions regarding the budget, domgross and intgross are heavily skewed to the right. we can again see what we identified before, that te budget for movies that passed the test is usually smaller and they have a lower imdb rating. In addition we identify a very high correlation between intgross and domgross, as expected. There is also a strong correlation between imdb rating and metascore but no correlation between budget and imdb rating or metascore.
## Categorical variables

```{r}
bechdel %>% 
  group_by(genre, test) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))
  
 
bechdel %>% 
  group_by(rated, test) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))
```
From the above tables we understand that most action, crime, mystery and animation movies fail the test but most horror movies pass the test. All documentaries and sci-fi movies fail the test but we have very few movies in these categories. On the other hand all musicals and thrillers pass the test, but with the same caviat, very small number of movies. All of the others genres are pretty evenly split.

Regarding the rating, all of them tend to fail the test but there are two notable ratings. NC-17 movies tend to fail the test the most (but we have a small number of movies in this rating) and PG-13 rated movies tend to have an even split.
# Train first models. `test ~ metascore + imdb_rating`

```{r}
lr_mod <- logistic_reg() %>% 
  set_engine(engine = "glm") %>% 
  set_mode("classification")

lr_mod


tree_mod <- decision_tree() %>% 
  set_engine(engine = "C5.0") %>% 
  set_mode("classification")

tree_mod 
```

```{r}


lr_fit <- lr_mod %>% # parsnip model
  fit(test ~ metascore + imdb_rating, # a formula
    data = bechdel_train # dataframe
  )

tree_fit <- tree_mod %>% # parsnip model
  fit(test ~ metascore + imdb_rating, # a formula
    data = bechdel_train # dataframe
  )
```

## Logistic regression

```{r}
lr_fit %>%
  broom::tidy()

lr_preds <- lr_fit %>%
  augment(new_data = bechdel_train) %>%
  mutate(.pred_match = if_else(test == .pred_class, 1, 0))

```

### Confusion matrix

```{r}
lr_preds %>% 
  conf_mat(truth = test, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")


```


## Decision Tree
```{r}
tree_preds <- tree_fit %>%
  augment(new_data = bechdel) %>%
  mutate(.pred_match = if_else(test == .pred_class, 1, 0)) 


```

```{r}
tree_preds %>% 
  conf_mat(truth = test, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")
```

## Draw the decision tree

```{r}
draw_tree <- 
    rpart::rpart(
        test ~ metascore + imdb_rating,
        data = bechdel_train, # uses data that contains both birth weight and `low`
        control = rpart::rpart.control(maxdepth = 5, cp = 0, minsplit = 10)
    ) %>% 
    partykit::as.party()
plot(draw_tree)

```

# Cross Validation

Run the code below. What does it return?

```{r}
set.seed(123)
bechdel_folds <- vfold_cv(data = bechdel_train, 
                          v = 10, 
                          strata = test)
bechdel_folds
```

## `fit_resamples()`

Trains and tests a resampled model.

```{r}
lr_fit <- lr_mod %>%
  fit_resamples(
    test ~ metascore + imdb_rating,
    resamples = bechdel_folds
  )


tree_fit <- tree_mod %>%
  fit_resamples(
    test ~ metascore + imdb_rating,
    resamples = bechdel_folds
  )
```


## `collect_metrics()`

Unnest the metrics column from a tidymodels `fit_resamples()`
```{r}

collect_metrics(lr_fit)
collect_metrics(tree_fit)


```


```{r}
tree_preds <- tree_mod %>% 
  fit_resamples(
    test ~ metascore + imdb_rating, 
    resamples = bechdel_folds,
    control = control_resamples(save_pred = TRUE) #<<
  )

# What does the data for ROC look like?
tree_preds %>% 
  collect_predictions() %>% 
  roc_curve(truth = test, .pred_Fail)  

# Draw the ROC
tree_preds %>% 
  collect_predictions() %>% 
  roc_curve(truth = test, .pred_Fail) %>% 
  autoplot()

```


# Build a better training set with `recipes`


## Collapse Some Categorical Levels

Do we have any `genre` with few observations?  Assign genres that have less than 3% to a new category 'Other'


```{r}
#| echo = FALSE
bechdel %>% 
  count(genre) %>% 
  mutate(genre = fct_reorder(genre, n)) %>% 
  ggplot(aes(x = genre, 
             y = n)) +
  geom_col(alpha = .8) +
  coord_flip() +
  labs(x = NULL) +
  geom_hline(yintercept = (nrow(bechdel_train)*.03), lty = 3)+
  theme_light()
```
Yes, Mystery, Fantasy, Sci-Fi, Thriller, Documentary and Musical.

```{r}
movie_rec <-
  recipe(test ~ .,
         data = bechdel_train) %>%
  
  # Genres with less than 3% will be in a category 'Other'
    step_other(genre, threshold = .03) 
```
  

## Before recipe

```{r}
#| echo = FALSE
bechdel_train %>% 
  count(genre, sort = TRUE)
```


## After recipe

```{r}
movie_rec %>% 
  prep() %>% 
  bake(new_data = bechdel_train) %>% 
  count(genre, sort = TRUE)
```

## `step_dummy()`

Converts nominal data into numeric dummy variables

```{r}
#| results = "hide"
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_dummy(all_nominal_predictors()) 

movie_rec 
```

## Let's think about the modelling 

What if there were no films with `rated` NC-17 in the training data?

-  Will the model have a coefficient for `rated` NC-17? 
No
-  What will happen if the test data includes a film with `rated` NC-17? 
We use step_novel in the recipe


```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% #Adds a catch-all level to a factor for any new values not encountered in model training
  step_zv(all_numeric(), -all_outcomes())  %>% #Intelligently handles zero variance variables
  step_normalize(all_numeric()) #Centers then scales numeric variable

movie_rec
```


# Define different models to fit

```{r}
## Model Building

# 1. Pick a `model type`
# 2. set the `engine`
# 3. Set the `mode`: regression or classification

# Logistic regression
log_spec <-  logistic_reg() %>%  # model type
  set_engine(engine = "glm") %>%  # model engine
  set_mode("classification") # model mode

# Show your model specification
log_spec

# Decision Tree
tree_spec <- decision_tree() %>%
  set_engine(engine = "C5.0") %>%
  set_mode("classification")

tree_spec

# Random Forest
library(ranger)

rf_spec <- 
  rand_forest() %>% 
  set_engine("ranger", importance = "impurity") %>% 
  set_mode("classification")


# Boosted tree (XGBoost)
library(xgboost)

xgb_spec <- 
  boost_tree() %>% 
  set_engine("xgboost") %>% 
  set_mode("classification") 

# K-nearest neighbour (k-NN)
knn_spec <- 
  nearest_neighbor(neighbors = 4) %>% # we can adjust the number of neighbors 
  set_engine("kknn") %>% 
  set_mode("classification") 
```


# Bundle recipe and model with `workflows`


```{r}
log_wflow <- # new workflow object
 workflow() %>% # use workflow function
 add_recipe(movie_rec) %>%   # use the new recipe
 add_model(log_spec)   # add your model spec

# show object
log_wflow


## A few more workflows

tree_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(tree_spec) 

rf_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(rf_spec) 

xgb_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(xgb_spec)

knn_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(knn_spec)

```

HEADS UP

1. How many models have you specified? 
5
2. What's the difference between a model specification and a workflow? 
A model specification focuses specifically on defining the structure and parameters of a single model, while a workflow refers to the broader sequence of steps involved in a data analysis project, encompassing data preprocessing, model building, evaluation, and other related tasks.
3. Do you need to add a formula (e.g., `test ~ .`)  if you have a recipe? 
No everything (except the folds) are already specified.


# Model Comparison

You now have all your models. Adapt the code from slides `code-from-slides-CA-housing.R`, line 400 onwards to assess which model gives you the best classification. 


```{r}
log_res <- log_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, accuracy,
      kap, roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)) 

# Show average performance over all folds (note that we use log_res):
log_res %>%  collect_metrics(summarize = TRUE)

# Show performance for every single fold:
log_res %>%  collect_metrics(summarize = FALSE)

## Decision Tree results

tree_res <-
  tree_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

## Random Forest

rf_res <-
  rf_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

## Boosted tree - XGBoost

xgb_res <- 
  xgb_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

## K-nearest neighbour

knn_res <- 
  knn_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

## Model Comparison

log_metrics <- 
  log_res %>% 
  collect_metrics(summarise = TRUE) %>%
  # add the name of the model to every row
  mutate(model = "Logistic Regression") 

tree_metrics <- 
  tree_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Decision Tree")

rf_metrics <- 
  rf_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Random Forest")

xgb_metrics <- 
  xgb_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "XGBoost")

knn_metrics <- 
  knn_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Knn")

# create dataframe with all models
model_compare <- bind_rows(log_metrics,
                           tree_metrics,
                           rf_metrics,
                           xgb_metrics,
                           knn_metrics) 

#Pivot wider to create barplot
  model_comp <- model_compare %>% 
  select(model, .metric, mean, std_err) %>% 
  pivot_wider(names_from = .metric, values_from = c(mean, std_err)) 

# show mean are under the curve (ROC-AUC) for every model
model_comp %>% 
  arrange(mean_roc_auc) %>% 
  mutate(model = fct_reorder(model, mean_roc_auc)) %>% # order results
  ggplot(aes(model, mean_roc_auc, fill=model)) +
  geom_col() +
  coord_flip() +
  scale_fill_brewer(palette = "Blues") +
   geom_text(
     size = 3,
     aes(label = round(mean_roc_auc, 2), 
         y = mean_roc_auc + 0.08),
     vjust = 1
  )+
  theme_light()+
  theme(legend.position = "none")+
  labs(y = NULL)
```



```{r}
# - `last_fit()`  fits a model to the whole training data and evaluates it on the test set. 
# - provide the workflow object of the best model as well as the data split object (not the training data). 
last_fit_xgb <- last_fit(xgb_wflow, 
                        split = data_split,
                        metrics = metric_set(
                          accuracy, f_meas, kap, precision,
                          recall, roc_auc, sens, spec))

last_fit_xgb %>% collect_metrics(summarize = TRUE)

#Compare to training
xgb_res %>% collect_metrics(summarize = TRUE)


## Variable importance using `{vip}` package

library(vip)

last_fit_xgb %>% 
  pluck(".workflow", 1) %>%   
  pull_workflow_fit() %>% 
  vip(num_features = 10) +
  theme_light()


## Final Confusion Matrix

last_fit_xgb %>%
  collect_predictions() %>% 
  conf_mat(test, .pred_class) %>% 
  autoplot(type = "heatmap")


## Final ROC curve
last_fit_xgb %>% 
  collect_predictions() %>% 
  roc_curve(test, .pred_Fail) %>% 
  autoplot()
```


# Deliverables

There is a lot of explanatory text, comments, etc. You do not need these, so delete them and produce a stand-alone document that you could share with someone. Knit the edited and completed R Markdown (Rmd) file as a Word or HTML document (use the "Knit" button at the top of the script editor window) and upload it to Canvas. You must be commiting and pushing your changes to your own Github repo as you go along.

# Details

-   Who did you collaborate with: nobody
-   Approximately how much time did you spend on this problem set: 3
-   What, if anything, gave you the most trouble: 

**Please seek out help when you need it,** and remember the [15-minute rule](https://dsb2023.netlify.app/syllabus/#the-15-minute-rule){target="_blank"}. You know enough R (and have enough examples of code from class and your readings) to be able to do this. If you get stuck, ask for help from others, post a question on Slack-- and remember that I am here to help too!

> As a true test to yourself, do you understand the code you submitted and are you able to explain it to someone else?

# Rubric

13/13: Problem set is 100% completed. Every question was attempted and answered, and most answers are correct. Code is well-documented (both self-documented and with additional comments as necessary). Used tidyverse, instead of base R. Graphs and tables are properly labelled. Analysis is clear and easy to follow, either because graphs are labeled clearly or you've written additional text to describe how you interpret the output. Multiple Github commits. Work is exceptional. I will not assign these often.

8/13: Problem set is 60--80% complete and most answers are correct. This is the expected level of performance. Solid effort. Hits all the elements. No clear mistakes. Easy to follow (both the code and the output). A few Github commits.

5/13: Problem set is less than 60% complete and/or most answers are incorrect. This indicates that you need to improve next time. I will hopefully not assign these often. Displays minimal effort. Doesn't complete all components. Code is poorly written and not documented. Uses the same type of plot for each graph, or doesn't use plots appropriate for the variables being analyzed. No Github commits.
